---
title: "Rethinking_Stats_Chapter_14.Rmd"
author: "Alex Bayers"
date: "26/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Easy
### Problem 1
Our base model is as follows:

$$
\\T_i \sim  Poisson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_i
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1)
$$
The model with measurement error is as follows:

$$
\\T_i \sim  Posson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_{EST, i}
\\P_{OBS, i} \sim Normal(P_{EST, i}, P_{SE,i})
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1)
$$

### Problem 2

Our new model for unobserved populations follows below, allowing for unobserved populations.

$$
\\T_i \sim  Poisson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_i
\\P_i \sim Normal(\nu, \sigma)
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1),
\\nu \sim Normal(34000, 30000)
\\\sigma \sim Cauchy(10000)
$$

## Medium
### Problem 1
In the mathematical model, when we have data missing, we are assuming a given distribution for the missing data.  In this case, we assume a mean for the missing data $$ \nu $$, and a standard deviation $$ \sigma_n $$.

### Problem 2
```{r}
```
