---
title: "Rethinking_Stats_Chapter_14.Rmd"
author: "Alex Bayers"
date: "26/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Easy
### Problem 1
Our base model is as follows:

$$
\\T_i \sim  Poisson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_i
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1)
$$
The model with measurement error is as follows:

$$
\\T_i \sim  Posson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_{EST, i}
\\P_{OBS, i} \sim Normal(P_{EST, i}, P_{SE,i})
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1)
$$

### Problem 2

Our new model for unobserved populations follows below, allowing for unobserved populations.

$$
\\T_i \sim  Poisson(\mu_i)
\\log \> \mu_i = \alpha \> + \> \beta * log \> P_i
\\P_i \sim Normal(\nu, \sigma)
\\\alpha \sim Normal(0, 10)
\\\beta \sim Nornal(0, 1),
\\nu \sim Normal(34000, 30000)
\\\sigma \sim Cauchy(10000)
$$

## Medium
### Problem 1
In the mathematical model, when we have data missing, we are assuming a given distribution for the missing data.  In this case, we assume a mean for the missing data $$ \nu $$, and a standard deviation $$ \sigma_n $$.

### Problem 2
```{r}
## R code 6.21
data(milk)
d <- milk
d$neocortex.prop <- d$neocortex.perc / 100
data_list <- list(
  kcal = d$kcal.per.g,
  neocortex = d$neocortex.prop,
  logmass = log(d$mass)
)

m6.11 <- map2stan(
    alist(
        kcal ~ dnorm(mu, sigma),
        mu ~ dnorm(0, 1),
        sigma ~ dcauchy(0, 1),
        neocortex ~ dnorm(nu, sigma_N),
        nu ~ dnorm(0.5, 2),
        sigma_N ~ dcauchy(0, 2)
    ),
    data=data_list,
    iter = 5000,
    warmup = 1000,
    cores = 4,
    chains = 4)

precis(m6.11, depth = 2)
```

```{r}
m6.12 <- map2stan(
    alist(
        kcal ~ dnorm(mu, sigma),
        mu <- a  + bn*neocortex,
        a ~ dnorm(0, 1),
        bn ~ dnorm(0, 10),
        sigma ~ dcauchy(0, 1),
        neocortex ~ dnorm(nu, sigma_N),
        nu ~ dnorm(0.5, 2),
        sigma_N ~ dcauchy(0, 1)
    ),
    data=data_list,
    iter = 5000,
    warmup = 1000,
    cores = 4,
    chains = 4)

precis(m6.12, depth = 2)
```

```{r}
m6.13 <- map2stan(
    alist(
        kcal ~ dnorm(mu, sigma),
        mu <- a  + bm*logmass,
        a ~ dnorm(0, 1),
        bm ~ dnorm(0, 10),
        sigma ~ dcauchy(0, 1),
        neocortex ~ dnorm(nu, sigma_N),
        nu ~ dnorm(0.5, 2),
        sigma_N ~ dcauchy(0, 1)
    ),
    data=data_list,
    iter = 5000,
    warmup = 1000,
    cores = 1,
    chains =14)

precis(m6.13, depth = 2)
```

```{r}

m6.14 <- map2stan(
    alist(
        kcal ~ dnorm(mu, sigma),
        mu <- a  + bn*neocortex + bm*logmass,
        a ~ dnorm(0, 1),
        bm ~ dnorm(0, 10),
        bn ~ dnorm(0, 10),
        sigma ~ dcauchy(0, 1),
        neocortex ~ dnorm(nu, sigma_N),
        nu ~ dnorm(0.5, 1),
        sigma_N ~ dcauchy(0, 1)
    ),
    data=data_list,
    iter = 5000,
    warmup = 1000,
    cores = 1,
    chains =14)

precis(m6.14, depth = 2)
```

```{r}

m6.15 <- map2stan(
    alist(
        kcal ~ dnorm(mu, sigma),
        mu <- a  + bn*neocortex + bm*logmass,
        a ~ dnorm(0, 1),
        bm ~ dnorm(0, 10),
        bn ~ dnorm(0, 10),
        sigma ~ dcauchy(0, 1),
        neocortex ~ dnorm(nu, sigma_N),
        nu ~ a_N + gM*logmass,
        a_N ~ dnorm(0.5, 1),
        gM ~ dnorm(0, 10),
        sigma_N ~ dcauchy(0, 1)
    ),
    data=data_list,
    iter = 5000,
    warmup = 1000,
    cores = 1,
    chains =14)

precis(m6.15, depth = 2)
```

```{r}
compare(m6.12, m6.13, m6.14, m6.15)
```
The multiple imputation model fits similar to the model with the complete cases.  

```{r}
## R code 6.21
data(milk)
d <- milk[ complete.cases(milk) , ]
d$neocortex <- d$neocortex.perc / 100
dim(d)

## R code 6.22
a.start <- mean(d$kcal.per.g)
sigma.start <- log(sd(d$kcal.per.g))
m6.11 <- map(
    alist(
        kcal.per.g ~ dnorm( a , exp(log.sigma) )
    ) ,
    data=d , start=list(a=a.start,log.sigma=sigma.start) )
m6.12 <- map(
    alist(
        kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
        mu <- a + bn*neocortex
    ) ,
    data=d , start=list(a=a.start,bn=0,log.sigma=sigma.start) )
m6.13 <- map(
    alist(
        kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
        mu <- a + bm*log(mass)
    ) ,
    data=d , start=list(a=a.start,bm=0,log.sigma=sigma.start) )
m6.14 <- map(
    alist(
        kcal.per.g ~ dnorm( mu , exp(log.sigma) ) ,
        mu <- a + bn*neocortex + bm*log(mass)
    ) ,
    data=d , start=list(a=a.start,bn=0,bm=0,log.sigma=sigma.start) )

precis(m6.14)
```

The coefficient estimates are similar, though all shrunken.

###Problem 3
```{r}
## R code 14.2
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

## R code 14.3
dlist <- list(
    div_obs=d$Divorce,
    div_sd=d$Divorce.SE,
    R=d$Marriage,
    A=d$MedianAgeMarriage
)

m14.1 <- map2stan(
    alist(
        div_est ~ dnorm(mu,sigma),
        mu <- a + bA*A + bR*R,
        div_obs ~ dnorm(div_est,div_sd),
        a ~ dnorm(0,10),
        bA ~ dnorm(0,10),
        bR ~ dnorm(0,10),
        sigma ~ dcauchy(0,2.5)
    ) ,
    data=dlist ,
    start=list(div_est=dlist$div_obs) ,
    WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
    control=list(adapt_delta=0.95) )

## R code 14.4
precis( m14.1 , depth=2 )

## R code 14.3
new_dlist <- list(
    div_obs=d$Divorce,
    div_sd=d$Divorce.SE*2,
    R=d$Marriage,
    A=d$MedianAgeMarriage
)

m14.2 <- map2stan(
    alist(
        div_est ~ dnorm(mu,sigma),
        mu <- a + bA*A + bR*R,
        div_obs ~ dnorm(div_est,div_sd),
        a ~ dnorm(0,10),
        bA ~ dnorm(0,10),
        bR ~ dnorm(0,10),
        sigma ~ dcauchy(0,2.5)
    ) ,
    data=new_dlist ,
    start=list(div_est=new_dlist$div_obs) ,
    WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
    control=list(adapt_delta=0.95) )

coeftab(m14.1, m14.2)
```

When we increase the variance on the measurement error, the coefficient bR goes up by a factor of 3, while the standard deviation goes down by 2/3.